---
title: 'SQL: GROUP BY Statements - Aggregate Functions'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: GROUP BY Statements

### Aggregate Functions

The Aggregate functions combine all the rows into a single row of data.

#### AVG
The AVG function returns the combined average of rows.

**Syntax**: 

```{sql, eval = FALSE}
-- AVG
SELECT AVG(column_1) 
FROM table_name;
```


**Example**: Obtain the average amount from the payment table

```{sql, eval = FALSE}
-- Obtain the average amount from the payment table
SELECT AVG(amount)
FROM payment;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT AVG(amount)
FROM payment
                
")
```

#### MIN
The MIN function returns the minimum value of the rows. 

**Syntax**: 

```{sql, eval = FALSE}
-- MIN
SELECT MIN(column_1) 
FROM table_name;
```


**Example**: Obtain the minimum amount from the payment table

```{sql, eval = FALSE}
-- Obtain the minimum amount from the payment table
SELECT MIN(amount)
FROM payment;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT MIN(amount)
FROM payment
                
")
```

#### MAX
The MAX function returns the maximum value of the rows.

**Syntax**: 

```{sql, eval = FALSE}
-- MAX
SELECT MAX(column_1) 
FROM table_name;
```


**Example**: Obtain the maximum amount from the payment table

```{sql, eval = FALSE}
-- Obtain the minimum amount from the payment table
SELECT MAX(amount)
FROM payment;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT MAX(amount)
FROM payment
                
")
```


#### SUM
The SUM function returns the combined sum of rows. 

**Syntax**: 

```{sql, eval = FALSE}
-- MAX
SELECT MAX(column_1) 
FROM table_name;
```


**Example**: Obtain the total amount from the payment table

```{sql, eval = FALSE}
-- Obtain the total amount from the payment table
SELECT SUM(amount)
FROM payment;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT SUM(amount)
FROM payment
                
")
```