---
title: 'SQL: GROUP BY Statements - GROUP BY'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: GROUP BY Statements


### GROUP BY

The GROUP BY clause divides thw row return from the SELECT statement into groups.

**Syntax**: 

```{sql, eval = FALSE}
-- GROUP BY
SELECT column_1, aggregate_function(column_2)
FROM table_name
GROUP BY column_1;
```


**Example**: GROUP BY without an aggregrate function

```{sql, eval = FALSE}
-- GROUP BY without an aggregrate function
SELECT customer_id
FROM payment
GROUP BY customer_id;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT customer_id
FROM payment
GROUP BY customer_id
                
")
```

**Example**: GROUP BY with SUM function

```{sql, eval = FALSE}
-- GROUP BY with SUM function
SELECT customer_id, sum(AMOUNT)
FROM payment
GROUP BY customer_id;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT customer_id, sum(AMOUNT)
FROM payment
GROUP BY customer_id
                
")
```

**Example**: GROUP BY with SUM function, and now use ORDER BY in a descending order

```{sql, eval = FALSE}
-- GROUP BY with SUM function, and now use ORDER BY in a descending order
SELECT customer_id, sum(AMOUNT)
FROM payment
GROUP BY customer_id
ORDER BY SUM(amount) DESC;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT customer_id, sum(AMOUNT)
FROM payment
GROUP BY customer_id
ORDER BY SUM(amount) DESC
                
")
```

**Example**: Count the number of transactions each staff member has been processing

```{sql, eval = FALSE}
-- Count the number of transactions each staff member has been processing
SELECT staff_id, COUNT(payment_id)
FROM payment
GROUP BY staff_id;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT staff_id, COUNT(payment_id)
FROM payment
GROUP BY staff_id
                
")
```

**Example**: How many films of rating type do we have?

```{sql, eval = FALSE}
-- Count the number of films for each raing type
SELECT rating, COUNT(rating) 
FROM FILM
GROUP BY rating;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT rating, COUNT(rating) 
FROM FILM
GROUP BY rating
                
")
```

**Example**: How many rental durations are in the film table.

```{sql, eval = FALSE}
-- Count the number of rental durations
SELECT rental_duration, COUNT(rental_duration)
FROM FILM
GROUP BY rental_duration;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT rental_duration, COUNT(rental_duration)
FROM FILM
GROUP BY rental_duration
                
")
```


**Example**: What is the average rental rate for a movie's rating?

```{sql, eval = FALSE}
-- Average rental rate for a movie's rating
SELECT rating, AVG(rental_rate)
FROM film
GROUP BY rating;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT rating, AVG(rental_rate)
FROM film
GROUP BY rating
                
")
```

### Challenge

**Problem**: How many payments did each staff member handle? And how much was the total amount processed by each staff member?

```{sql, eval = FALSE}
-- Count the number of payments for each staff member and obtain the total amount
SELECT staff_id, COUNT(amount), SUM(amount) 
FROM payment
GROUP BY staff_id;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT staff_id, COUNT(amount), SUM(amount) 
FROM payment
GROUP BY staff_id;
                
")
```

**Problem**: What is the average replacement cost of movies by rating?

```{sql, eval = FALSE}
-- Average replacement cost of movies by rating
SELECT rating, AVG(replacement_cost)
FROM film
GROUP BY rating;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT rating, AVG(replacement_cost)
FROM film
GROUP BY rating
                
")
```

**Problem**: Obtain the customer ids of the top 5 spenders

```{sql, eval = FALSE}
-- Customer IDs of the top 5 spenders
SELECT customer_id, SUM(amount)
FROM payment
GROUP BY customer_id
ORDER BY SUM(amount) DESC
LIMIT 5;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT customer_id, SUM(amount)
FROM payment
GROUP BY customer_id
ORDER BY SUM(amount) DESC
LIMIT 5;
                
")
```