---
title: 'SQL: Advanced SQL Commands - Timestamps and Extract'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: JOINS - Advanced SQL Commands - Timestamps and Extract


### Timestamps 

Timestamp data types are used to retain information.

### Extract

The extract function extracts parts from a date. Look at [extract functions](https://www.postgresql.org/docs/8.1/functions-datetime.html) for a more detailed list on extract functions.

**Syntax**: 

```{sql, eval = FALSE}
-- EXTRACT
extract (unit_from_date)
```


**Example**: Examine the payment_date in the payment table

```{sql, eval = FALSE}
-- Examine the payment_date in the payment table. Extract the day from the payment date
SELECT customer_id, extract(day from payment_date) AS day
FROM
payment
limit 10;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT customer_id, extract(day from payment_date) AS day
FROM
payment
limit 10
")
```


**Example**: What is the total amount of expenditure by the month?
```{sql, eval = FALSE}
-- What is the total amount of expenditure by the month?
SELECT SUM(amount), extract(month from payment_date) AS month
FROM payment
GROUP BY month
ORDER BY SUM(amount) DESC ;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT SUM(amount), extract(month from payment_date) AS month
FROM payment
GROUP BY month
ORDER BY SUM(amount) DESC
")
```
