---
title: 'SQL: JOINS - OUTER JOIN'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: JOINS - OUTER JOIN


### OUTER JOINs

#### FULL OUTER JOIN

The FULL OUTER JOIN produces the set of all records in Table A and Table B, with matching records from both sides where available. If there is no match, the missing side will contain null.


**Syntax**: 

```{sql, eval = FALSE}
-- FULL OUTER JOIN
SELECT *
FROM TableA
FULL OUTER JOIN TableB ON TableA.name = TableB.name;
```

#### FULL OUTER JOIN with WHERE

The FULL OUTER JOIN with WHERE produces the set of records unqiue to Table A and Table B, we perform the same full outer join, then exclude the records we don't want from both sides via a where clause.

**Syntax**: 

```{sql, eval = FALSE}
-- FULL OUTER JOIN
SELECT *
FROM TableA
LEFT OUTER JOIN TableB ON TableA.name = TableB.name
WHERE TableB.id IS null OR TableB.id IS null;
```

#### LEFT OUTER JOIN

The LEFT OUTER JOIN produces a complete set of records from Table A, with the matching records(where available) in Table B. If there is no match, the right side will contain null.

**Syntax**: 

```{sql, eval = FALSE}
-- FULL OUTER JOIN
SELECT *
FROM TableA
LEFT OUTER JOIN TableB ON TableA.name = TableB.name;
```




#### LEFT OUTER JOIN with WHERE

The LEFT OUTER JOIN with WHERE produces the set of records only in Table A, but not in Table B, we perform the same left outer join, then exclude the records we don't want from the right side via a where clause.

**Syntax**: 

```{sql, eval = FALSE}
-- FULL OUTER JOIN
SELECT *
FROM TableA
LEFT OUTER JOIN TableB ON TableA.name = TableB.name
WHERE TableB.id IS null;
```

### OUTER JOIN Example

**Example**: Perform a LEFT OUTHER JOIN with the film and inventory tables to obtain a copy of all movies that are not in iventory.

```{sql, eval = FALSE}
-- Perform a LEFT OUTHER JOIN with the film and inventory tables.
SELECT film.film_id, film.title, inventory_id
FROM film
LEFT OUTER JOIN inventory ON inventory.film_id = film.film_id
WHERE inventory.film_id IS NULL
ORDER BY film.film_id;
-- Second Solution
SELECT film.film_id, film.title, inventory_id
FROM film
LEFT OUTER JOIN inventory ON inventory.film_id = film.film_id
WHERE inventory_id is NULL
ORDER BY film.title;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT film.film_id, film.title, inventory_id
FROM film
LEFT OUTER JOIN inventory ON inventory.film_id = film.film_id
WHERE inventory.film_id IS NULL
ORDER BY film.film_id  
")
```


