---
title: 'SQL: JOINS - INNER JOIN'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: JOINS - INNER JOIN


### INNER JOIN

The INNER JOIN clause returns rows in A table that have the corresponding rows in the B table.


**Syntax**: 

```{sql, eval = FALSE}
-- INNER JOIN
SELECT A.pka, A.c1, B.pkb, B.c2
FROM A
INNER JOIN B ON A.pka = B.fka;
```


**Example**: Join the customer_id columns from the customer and payment tables

```{sql, eval = FALSE}
-- Join the customer_id columns from the customer and payment tables. Then, sort and have the customer_id = 2
SELECT customer.customer_id, first_name, last_name, email, amount, payment_date
FROM customer
INNER JOIN payment ON payment.customer_id = customer.customer_id
WHERE customer.customer_id = 2
ORDER BY customer.customer_id;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT customer.customer_id, first_name, last_name, email, amount, payment_date
FROM customer
INNER JOIN payment ON payment.customer_id = customer.customer_id
WHERE customer.customer_id = 2
ORDER BY customer.customer_id          
")
```


**Example**: Inner join the payment and staff tables

```{sql, eval = FALSE}
-- Inner join the payment and staff tables
SELECT payment_id, amount, first_name, last_name
FROM payment
INNER JOIN staff ON payment.staff_id = staff_staff_id;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT payment_id, amount, first_name, last_name
FROM payment
INNER JOIN staff ON payment.staff_id = staff_staff_id
")
```

**Example**: Inner join the inventory and film tables where the store id is 1. 

```{sql, eval = FALSE}
-- Inner join the inventory and film tables where store is 1
SELECT COUNT(title) AS copies_at_store1 
FROM inventory
INNER JOIN film ON inventory.film_id = film.film_id
WHERE store_id = 1
GROUP BY title 
ORDER BY title;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT COUNT(title) AS copies_at_store1 
FROM inventory
INNER JOIN film ON inventory.film_id = film.film_id
WHERE store_id = 1
GROUP BY title 
ORDER BY title;
")
```


**Example**: Inner join the film and language tables to present the language of the movie and film title.

```{sql, eval = FALSE}
-- Inner join the film and language tables to present the language of the movie and film title.
SELECT film.title, lan.name AS movie_language
FROM film
INNER JOIN language AS lan ON lan.language_id = film.language_id;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT film.title, lan.name AS movie_language
FROM film
INNER JOIN language AS lan ON lan.language_id = film.language_id
")
```