---
title: 'SQL: Statement Fundamentals - SELECT WHERE'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Statement Fundamentals


### SELECT WHERE

The SELECT DISTINCT function returns particular queries indicated in the rows and columns from a database.

**Conditions**:

| Operator       | Description           | 
| ------------- |:-------------:| 
| \(=\)         | Equal         | 
| \(>\)         | Greater than  | 
| \(<\)         | Less than     |
| \(\geq\)      | Greater than or equal to |
| \(\leq\)      | Less than or equal to |
| \(\neq\)      | Not equal |
| AND        | Logical operator AND |
| OR         | Logical operator OR |


**Syntax**:

```{sql, eval = FALSE}
-- Select particular rows in a database with specific conditions
SELECT column_1, column_2
FROM table_name
WHERE conditions;
```

**Example**: Select all customers whose first names are Jamie

```{sql, eval = FALSE}
-- Select all customers whose first names are Jamie
SELECT last_name, first_name
FROM customer
WHERE first_name = 'Jamie';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT last_name, first_name
FROM customer
WHERE first_name = 'Jamie'
")
```

**Example** : Select the customer whose first name is Jamie and last name is Rice

```{sql, eval = FALSE}
-- Select the customer whose first name is Jamie and last name is Rice
SELECT last_name, first_name
FROM customer
WHERE first_name = 'Jamie' AND last_name = 'Rice';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT last_name, first_name
FROM customer
WHERE first_name = 'Jamie' AND last_name = 'Rice'
")
```

**Example**: Who paid the rental with amount is either less than \$1 or greater than \$8

```{sql, eval = FALSE}
-- Select the customer who paid the rental with amount is either less than $1 or greater than $8
SELECT customer_id, amount, payment_date
FROM payment
WHERE amount <= 1 OR amount >= 8;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT customer_id, amount, payment_date
FROM payment
WHERE amount <= 1 OR amount >= 8
")
```

**Examples**: Various SELECT WHERE examples

```{sql, eval = FALSE}
-- Select all rows and columns from the customer table
SELECT *
FROM customer;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT *
FROM customer
")
```

```{sql, eval = FALSE}
-- Select queries from the customer table where the first name is Jared
SELECT * 
FROM customer
WHERE first_name = 'Jared';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT * 
FROM customer
WHERE first_name = 'Jared'
")
```

#### Challenge

**Problem**: What's the email for the customer with the name Nancy Thomas?

```{sql, eval = FALSE}
-- Select the email for the customer with the name Nancy Thomas
SELECT email
FROM customer
WHERE first_name = 'Nancy' AND last_name = 'Thomas';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT email
FROM customer
WHERE first_name = 'Nancy' AND last_name = 'Thomas'
")
```

**Problem**: Could you provide a description for the movie "Outlaw Hanky"?

```{sql, eval = FALSE}
-- Select the description for the movie "Outlaw Hanky"
SELECT description
FROM film
WHERE title = 'Outlaw Hanky';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT description
FROM film
WHERE title = 'Outlaw Hanky'
")
```

**Problem**: Could you provide the phone number for the customer who lives at "259 Ipoh Drive"?

```{sql, eval = FALSE}
-- Select the customer's phone number who lives at "259 Ipoh Drive"
SELECT phone
FROM address
WHERE address = '259 Ipoh Drive';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT phone
FROM address
WHERE address = '259 Ipoh Drive'
")
```

