---
title: 'SQL: Statement Fundamentals - ORDER BY'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Statement Fundamentals


### ORDER BY

Sorts the resulting set either in an ascending or descending order based on criteria chosen.

**Syntax**:

```{sql, eval = FALSE}
-- ORDER BY
SELECT column_1, column_2
FROM table_name
ORDER BY column_1 ASC/DESC;
```

**Example**: Sort the customers by first name in ascending order.

```{sql, eval = FALSE}
-- Sort by first_name in ascending order from the customer table
SELECT first_name, last_name
FROM customer
ORDER BY first_name ASC;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT first_name, last_name
FROM customer
ORDER BY first_name ASC
")
```

**Example**: Sort the customers by last name in descending order.

```{sql, eval = FALSE}
-- Sort by last_name in descending order from the customer table
SELECT first_name, last_name
FROM customer
ORDER BY last_name DESC;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT first_name, last_name
FROM customer
ORDER BY last_name DESC
")
```

**Example**: Sort the customers by first name in the ascending order first and then sort the sorted result set by the last name in descending order

```{sql, eval = FALSE}
-- Sort by first_name in the ascending order, and then sort by last name in descending order from the customer table
SELECT first_name, last_name
FROM customer
ORDER BY first_name ASC,
last_name DESC;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT first_name, last_name
FROM customer
ORDER BY first_name ASC,
last_name DESC
")
```

#### Challenge

**Problem**: Get the customer_ID numbers for the top 10 highest payment amounts. 

```{sql, eval = FALSE}
-- Select customer ID numbers and sort them based on the top 10 highest payment amounts
SELECT customer_ID, amount
FROM payment
ORDER BY amount DESC
LIMIT 10;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT customer_ID, amount
FROM payment
ORDER BY amount DESC
LIMIT 10
")
```

**Problem**: Get the titles of the movies with film ids 1-5.

```{sql, eval = FALSE}
-- Select the titles of the movies with film ids 1-5
SELECT title, film_id
FROM film
ORDER BY film_id ASC
LIMIT 5;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT title, film_id
FROM film
ORDER BY film_id ASC
LIMIT 5
")
```

