---
title: 'SQL: Statement Fundamentals - General Challenges'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Statement Fundamentals

### General Challenges


**Problem**: How many payment transactions were greater than $5.00?

```{sql, eval = FALSE}
--Return the number of payment transactions were greater than $5.00
SELECT COUNT(amount)
FROM payment
WHERE amount > 5;
```


```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT COUNT(amount)
FROM payment
WHERE amount > 5
")
```


**Problem**: How many actors have a first name that starts with the letter P

```{sql, eval = FALSE}
--Return the number of actors that have a first name which starts with the letter P
SELECT COUNT(actor)
FROM actor
WHERE first_name LIKE 'P%';
```



```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT COUNT(actor)
FROM actor
WHERE first_name LIKE 'P%'
")
```

**Problem**: How many unique districts are our customers from?

```{sql, eval = FALSE}
--Return the number of unique districts that are our customers from
SELECT COUNT(DISTINCT(district))
FROM address;
```



```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT COUNT(DISTINCT(district))
FROM address
")
```

**Problem**: Retrieve the list of names for those distinct districts from the previous question.

```{sql, eval = FALSE}
--Return the names of unique districts that are our customers from
SELECT DISTINCT district
FROM address;
```



```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT DISTINCT district
FROM address
")
```

**Problem**: How many films have a rating of R and a replacement cost between \$5 and \$15?

```{sql, eval = FALSE}
--Return the number of films have a rating of R and a replacement cost between $5 and $15
SELECT COUNT(*)
FROM film
WHERE rating = 'R' AND replacement_cost BETWEEN 5 AND 15;
```



```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT COUNT(*)
FROM film
WHERE rating = 'R' AND replacement_cost BETWEEN 5 AND 15
")
```


**Problem**: How many films have the word Truman somewhere in the title?

```{sql, eval = FALSE}
--Return the number of films have the word Truman somewhere in the title
SELECT COUNT(*)
FROM film
WHERE title LIKE '%Truman%';

```



```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,
"
SELECT COUNT(*)
FROM film
WHERE title LIKE '%Truman%'
")
```