---
title: 'SQL: Assessment Test 3 - Creating Databases and Tables'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Assessment Test 3 - Creating Databases and Tables


### Problem 

**Problem**: 

Create a new database called "School" this database should have two tables: teachers and students.

The students table should have columns for student_id, first_name,last_name, homeroom_number, phone,email, and graduation year.

The teachers table should have columns for teacher_id, first_name, last_name, homeroom_number, department, email, and phone.

The constraints are mostly up to you, but your table constraints do have to consider the following:

* We must have a phone number to contact students in case of an emergency.
* We must have ids as the primary key of the tables
* Phone numbers and emails must be unique to the individual.

Once you've made the tables, insert a student named Mark Watney (student_id=1) who has a phone number of 777-555-1234 and doesn't have an email. He graduates in 2035 and has 5 as a homeroom number.

Then insert a teacher names Jonas Salk (teacher_id = 1) who as a homeroom number of 5 and is from the Biology department. His contact info is: jsalk@school.org and a phone number of 777-555-4321.





```{sql, eval = FALSE}
-- Generate students
CREATE TABLE students(
student_id serial PRIMARY KEY,
first_name VARCHAR(50) NOT NULL,
last_name VARCHAR(50) NOT NULL,
homeroom_number integer,
phone VARCHAR(255) NOT NULL UNIQUE,
email VARCHAR(50) UNIQUE,
grad_year integer
);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="education", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE students(
student_id serial PRIMARY KEY,
first_name VARCHAR(50) NOT NULL,
last_name VARCHAR(50) NOT NULL,
homeroom_number integer,
phone VARCHAR(255) NOT NULL UNIQUE,
email VARCHAR(50) UNIQUE,
grad_year integer
)
                
")
```

```{sql, eval = FALSE}
-- Generate teachers
CREATE TABLE teachers(
teacher_id serial PRIMARY KEY,
first_name VARCHAR(50) NOT NULL,
last_name VARCHAR(50) NOT NULL,
homeroom_number integer,
department VARCHAR(50),
phone VARCHAR(255) UNIQUE,
email VARCHAR(50) UNIQUE
);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="education", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE teachers(
teacher_id serial PRIMARY KEY,
first_name VARCHAR(50) NOT NULL,
last_name VARCHAR(50) NOT NULL,
homeroom_number integer,
department VARCHAR(50),
phone VARCHAR(255) UNIQUE,
email VARCHAR(50) UNIQUE
)
                
")
```



```{sql, eval = FALSE}
-- INSERT students
INSERT INTO students (student_id, first_name, last_name, homeroom_number, phone, grad_year)
VALUES (1, 'Mark', 'Watney', 5, '777-555-1234', 2035);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="education", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
INSERT INTO students (student_id, first_name, last_name, homeroom_number, phone, grad_year)
VALUES (1, 'Mark', 'Watney', 5, '777-555-1234', 2035)
                
")
```


```{sql, eval = FALSE}
-- INSERT teachers
INSERT INTO teachers (teacher_id, first_name, last_name, homeroom_number, department, phone,  email)
VALUES (1, 'Jonas', 'Salk', 5, 'Biology',  '777-555-4321', 'jsalk@school.org');
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="education", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
INSERT INTO students (student_id, first_name, last_name, homeroom_number, phone, grad_year)
VALUES (1, 'Mark', 'Watney', 5, '777-555-1234', 2035)
                
")
```


