---
title: 'SQL: Assessment Test 1 - SQL Statement Fundamentals and GROUP BY Statements'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Assessment Test 1 - SQL Statement Fundamentals and GROUP BY Statements


### Problem 1

**Problem**: Return the customer IDs of customers who have spent at least $110 with the staff member who has an ID of 2.

```{sql, eval = FALSE}
-- Return the customer IDs of customers who have spent at least $110 with the staff member who has an ID of 2.
SELECT staff_id ,customer_id , SUM(amount)
FROM payment
WHERE staff_id = 2
GROUP BY customer_id, staff_id
HAVING SUM(amount) >= 110;
-- Solution TWO
SELECT customer_id, SUM(amount)
FROM payment
WHERE staff_id = 2
GROUP BY customer_ID
HAVING SUM(amount) >= 110; 
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT staff_id ,customer_id , SUM(amount)
FROM payment
WHERE staff_id = 2
GROUP BY customer_id, staff_id
HAVING SUM(amount) >= 110
                
")
```

### Problem 2

**Problem**: How many films begin with the letter J?

```{sql, eval = FALSE}
-- Count the number of films that begin with the letter J
SELECT COUNT(title)
FROM film
WHERE title LIKE 'J%';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT COUNT(title)
FROM film
WHERE title LIKE 'J%'
                
")
```


### Problem 3

**Problem**: What customer has the highest customer ID number whose name starts with an 'E' and has an address ID lower than 500?

```{sql, eval = FALSE}
-- What customer has the highest customer ID number whose name starts with an 'E' and has an address ID lower than 500?
SELECT first_name, last_name, customer_id 
FROM customer
WHERE first_name LIKE 'E%' AND address_id < 500
ORDER BY customer_id DESC
LIMIT 1;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT first_name, last_name, customer_id 
FROM customer
WHERE first_name LIKE 'E%' AND address_id < 500
ORDER BY customer_id DESC
LIMIT 1
                
")
```