---
title: 'SQL: Assessment Test 2 - JOINS and Advanced SQL Commands'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Assessment Test 2 - JOINS and Advanced SQL Commands


### Problem 1

**Problem**: How can you retrieve all the information from the cd.facilities table?

```{sql, eval = FALSE}
-- Retrieve all the information from the cd.facilities table?
SELECT *
FROM cd.facilities;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT *
FROM cd.facilities
                
")
```

### Problem 2

**Problem**: You want to print out a list of all of the facilities and their cost to members. How would you retrieve a list of only facility names and costs?

```{sql, eval = FALSE}
-- Retrieve a list of only facility names and costs?
SELECT name, membercost
FROM cd.facilities;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT name, membercost
FROM cd.facilities
                
")
```


### Problem 3

**Problem**: How can you produce a list of facilities that charge a fee to members?

```{sql, eval = FALSE}
-- Produce a list of facilities that charge a fee to members
SELECT name, membercost
FROM cd.facilities
WHERE membercost > 0;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT name, membercost
FROM cd.facilities
WHERE membercost > 0
                
")
```

### Problem 4

**Problem**: How can you produce a list of facilities that charge a fee to members, and that fee is less than 1/50th of the monthly maintenance cost? Return the facid, facility name, member cost, and monthly maintenance of the facilities in question.

```{sql, eval = FALSE}
-- Produce a list of facilities that charge a fee to members, and that fee is less than 1/50th of the monthly maintenance cost. Return the facid, facility name, member cost, and monthly maintenance of the facilities in question.
SELECT facid, name, membercost, monthlymaintenance
FROM cd.facilities
WHERE membercost > 0 AND membercost < monthlymaintenance/50;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT facid, name, membercost, monthlymaintenance
FROM cd.facilities
WHERE membercost > 0 AND membercost < monthlymaintenance/50
                
")
```


### Problem 5

**Problem**: How can you produce a list of all facilities with the word 'Tennis' in their name?

```{sql, eval = FALSE}
-- Produce a list of all facilities with the word 'Tennis' in their name
SELECT name
FROM cd.facilities
WHERE name LIKE '%Tennis%';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT name
FROM cd.facilities
WHERE name LIKE '%Tennis%'
                
")
```

### Problem 6

**Problem**: How can you retrieve the details of facilities with ID 1 and 5? Try to do it without using the OR operator.

```{sql, eval = FALSE}
-- Retrieve the details of facilities with ID 1 and 5? Try to do it without using the OR operator.
SELECT facid, name
FROM cd.facilities
WHERE facid IN (1,5);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT facid, name
FROM cd.facilities
WHERE facid IN (1,5)
                
")
```

### Problem 7

**Problem**: How can you produce a list of members who joined after the start of September 2012? Return the memid, surname, firstname, and joindate of the members in question.

```{sql, eval = FALSE}
-- Produce a list of members who joined after the start of September 2012. Return the memid, surname, firstname, and joindate of the members in question.
SELECT *
FROM cd.members
WHERE joindate >= '2012-09-01';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT *
FROM cd.members
WHERE joindate >= '2012-09-01
                
")
```

### Problem 8

**Problem**: How can you produce an ordered list of the first 10 surnames in the members table? The list must not contain duplicates.

```{sql, eval = FALSE}
-- Produce an ordered list of the first 10 surnames in the members table? The list must not contain duplicates.
SELECT DISTINCT surname
FROM cd.members
ORDER BY surname ASC
LIMIT 10;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT DISTINCT surname
FROM cd.members
ORDER BY surname ASC
LIMIT 10
                
")
```

### Problem 9

**Problem**: You'd like to get the signup date of your last member. How can you retrieve this information?

```{sql, eval = FALSE}
-- You'd like to get the signup date of your last member. How can you retrieve this information?
SELECT max(joindate)
FROM cd.members;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT max(joindate)
FROM cd.members
                
")
```

### Problem 10

**Problem**: Produce a count of the number of facilities that have a cost to guests of 10 or more.

```{sql, eval = FALSE}
-- Produce a count of the number of facilities that have a cost to guests of 10 or more.
SELECT COUNT(*)
FROM cd.facilities
WHERE guestcost >= 10;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT COUNT(*)
FROM cd.facilities
WHERE guestcost >= 10
                
")
```

### Problem 12

**Problem**: Produce a list of the total number of slots booked per facility in the month of September 2012. Produce an output table consisting of facility id and slots, sorted by the number of slots.

```{sql, eval = FALSE}
-- Produce a list of the total number of slots booked per facility in the month of September 2012. Produce an output table consisting of facility id and slots, sorted by the number of slots.
SELECT facid, SUM(slots) AS Total_Number_of_Slots
FROM cd.bookings
WHERE cd.bookings.starttime BETWEEN '2012-09-01' AND '2012-09-30'
GROUP by facid
ORDER BY SUM(slots);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT facid, SUM(slots) AS Total_Number_of_Slots
FROM cd.bookings
WHERE cd.bookings.starttime BETWEEN '2012-09-01' AND '2012-09-30'
GROUP by facid
ORDER BY SUM(slots)
")
```


### Problem 13

**Problem**: Produce a list of facilities with more than 1000 slots booked. Produce an output table consisting of facility id and total slots, sorted by facility id.

```{sql, eval = FALSE}
-- Produce a list of facilities with more than 1000 slots booked. Produce an output table consisting of facility id and total slots, sorted by facility id.
SELECT facid, SUM(slots) AS Total_Number_of_Slots
FROM cd.bookings
GROUP BY facid
HAVING SUM(slots) > 1000
ORDER BY facid;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT facid, SUM(slots) AS Total_Number_of_Slots
FROM cd.bookings
GROUP BY facid
HAVING SUM(slots) > 1000
ORDER BY facid
")
```


### Problem 14

**Problem**: How can you produce a list of the start times for bookings for tennis courts, for the date '2012-09-21'? Return a list of start time and facility name pairings, ordered by the time.

```{sql, eval = FALSE}
-- Produce a list of facilities with more than 1000 slots booked. Produce an output table consisting of facility id and total slots, sorted by facility id.
SELECT starttime, name
FROM cd.bookings
INNER JOIN cd.facilities ON cd.facilities.facid = cd.bookings.facid
WHERE name LIKE '%Tennis Court%' AND starttime BETWEEN '2012-09-21 00:00:00' AND '2012-09-21 23:59:59' 
ORDER BY starttime;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT starttime, name
FROM cd.bookings
INNER JOIN cd.facilities ON cd.facilities.facid = cd.bookings.facid
WHERE name LIKE '%Tennis Court%' AND starttime BETWEEN '2012-09-21 00:00:00' AND '2012-09-21 23:59:59' 
ORDER BY starttime
")
```

### Problem 15

**Problem**: How can you produce a list of the start times for bookings by members named 'David Farrell'?

```{sql, eval = FALSE}
-- Produce a list of the start times for bookings by members named 'David Farrell'?
SELECT starttime
FROM cd.bookings
INNER JOIN cd.members ON cd.members.memid = cd.bookings.memid
WHERE surname LIKE '%Farrell%' AND firstname LIKE '%David%'  
ORDER BY starttime;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="exercises", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT starttime
FROM cd.bookings
INNER JOIN cd.members ON cd.members.memid = cd.bookings.memid
WHERE surname LIKE '%Farrell%' AND firstname LIKE '%David%'  
ORDER BY starttime;
")
```