---
title: 'SQL: Creating Databases and Tables - Create Table Code'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Creating Databases and Tables - Create Table Code


### CREATE TABLE

The CREATE TABLE function creates a table in a database.

Postgre SQL Column Constraints

* NOT NULL - the value of the column cannot be NULL

* UNIQUE - the value of the column must be unique across the whole table

* Primary Key - this constraint is the combination of NOT Null and UNIQUE constraints

* CHECK - enables to check a condition when you insert or update data

* REFERENCES - constrains the value of the column that exists in a column in another table (Define the foreign key)


PostgreSQL Table Constraints

* Unique (column_list) - to force the value stored in the columns listed inside the parentheses to be unique

* Primary Key (column_list) - to define the primary key that consists of multiple column

* Check (condition) - to check a condition when inserting or updating data

* References - to constrain the value store din the column that must exist in a column in another table



**Syntax**: 

```{sql, eval = FALSE}
-- CREATE TABLE
CREATE TABLE table_name
(column_name TYPE column_constraint,
table_constraint)
INHERITS existing_table_name;
```


**Example**: Account table

```{sql, eval = FALSE}
-- Account table
CREATE TABLE account(
user_id serial PRIMARY KEY,
username VARCHAR (50) UNIQUE NOT NULL,
password VARCHAR (50) NOT NULL,
email VARCHAR (355) UNIQUE NOT NULL,
created_on TIMESTAMP NOT NULL,
last_login TIMESTAMP
);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE account(
user_id serial PRIMARY KEY,
username VARCHAR (50) UNIQUE NOT NULL,
password VARCHAR (50) NOT NULL,
email VARCHAR (355) UNIQUE NOT NULL,
created_on TIMESTAMP NOT NULL,
last_login TIMESTAMP
)
")
```

**Example**: Role table

```{sql, eval = FALSE}
-- Role table.
CREATE TABLE role(
role_id serial PRIMARY KEY,
role_name VARCHAR (255) UNIQUE NOT NULL
);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE role(
role_id serial PRIMARY KEY,
role_name VARCHAR (255) UNIQUE NOT NULL
)
")
```

**Example**: Account-role table

```{sql, eval = FALSE}
-- Account-role table
CREATE TABLE account_role
(
  user_id integer NOT NULL,
  role_id integer NOT NULL,
  grant_date timestamp without time zone,
  PRIMARY KEY (user_id, role_id),
  CONSTRAINT account_role_role_id_fkey FOREIGN KEY (role_id)
      REFERENCES role (role_id) MATCH SIMPLE
      ON UPDATE NO ACTION ON DELETE NO ACTION,
  CONSTRAINT account_role_user_id_fkey FOREIGN KEY (user_id)
      REFERENCES account (user_id) MATCH SIMPLE
      ON UPDATE NO ACTION ON DELETE NO ACTION
);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE account_role
(
  user_id integer NOT NULL,
  role_id integer NOT NULL,
  grant_date timestamp without time zone,
  PRIMARY KEY (user_id, role_id),
  CONSTRAINT account_role_role_id_fkey FOREIGN KEY (role_id)
      REFERENCES role (role_id) MATCH SIMPLE
      ON UPDATE NO ACTION ON DELETE NO ACTION,
  CONSTRAINT account_role_user_id_fkey FOREIGN KEY (user_id)
      REFERENCES account (user_id) MATCH SIMPLE
      ON UPDATE NO ACTION ON DELETE NO ACTION
)
")
```

### Challenge

**Problem** : A customer's first name, last name,email,sign-up date, and number of minutes spent on the dvd rental site. You should also have some sort of id tracker for them. 

```{sql, eval = FALSE}
-- Role table.
CREATE TABLE potential_leads(
user_id serial PRIMARY KEY, 
first_name VARCHAR (50) NOT NULL,
last_name VARCHAR (50) NOT NULL,
email VARCHAR (355) UNIQUE NOT NULL,
sign_up_date TIMESTAMP NOT NULL,
minutes_spent_on_dvd_rental_site integer NOT NULL
);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="dvdrental", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE potential_leads(
user_id serial PRIMARY KEY, 
first_name VARCHAR (50) NOT NULL,
last_name VARCHAR (50) NOT NULL,
email VARCHAR (355) UNIQUE NOT NULL,
sign_up_date TIMESTAMP NOT NULL,
minutes_spent_on_dvd_rental_site integer NOT NULL
)
")
```
