---
title: 'SQL: Creating Databases and Tables - Insert'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Creating Databases and Tables - Insert


### INSERT

When you create a new table, it does not have any data. The first thing you often do is to insert new rows into the table. 

SQL provides the INSERT statement that allows you to insert one or more rows into a table at a time




**Syntax**: 

```{sql, eval = FALSE}
-- INSERT 
INSERT INTO table(column1, column2, ...)
VALUES (value1, value2, ...);
-- INSERT data that comes from another table
INSERT INTO table
SELECT column1, column2, ...
FROM another_table
WHERE condition;
```


**Example**: Link

```{sql, eval = FALSE}
-- Create Table
CREATE TABLE link(
ID serial PRIMARY KEY,
url VARCHAR(255) NOT NULL,
name VARCHAR(255) NOT NULL,
description VARCHAR(255),
rel VARCHAR (50)
);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE link(
ID serial PRIMARY KEY,
url VARCHAR(255) NOT NULL,
name VARCHAR(255) NOT NULL,
description VARCHAR(255),
rel VARCHAR (50)
)
")
```


```{sql, eval = FALSE}
-- Insert Google
INSERT INTO link(url,name)
VALUES ('www.google.com', 'Google');
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
INSERT INTO link(url,name)
VALUES ('www.google.com', 'Google')
")
```

**Example**: Link #2 Insert

```{sql, eval = FALSE}
-- INSERT Yahoo
INSERT INTO link(url,name)
VALUES ('www.yahoo.com', 'Yahoo');
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
INSERT INTO link(url,name)
VALUES ('www.yahoo.com', 'Yahoo')
")
```

**Example**: Link #3 Insert Multiple Values

```{sql, eval = FALSE}
-- INSERT multiple values
INSERT INTO link(url,name)
VALUES ('www.bing.com', 'Bing'), ('www.amazon.com', 'Amazon');
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
INSERT INTO link(url,name)
VALUES ('www.bing.com', 'Bing'), ('www.amazon.com', 'Amazon')
")
```


```{sql, eval = FALSE}
-- Select
SELECT * FROM link;

```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT * FROM link
")
```


**Example**: Insert another table

```{sql, eval = FALSE}
-- Create Table similar to link
CREATE TABLE link_copy (LIKE link);
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
CREATE TABLE link_copy (LIKE link)
")
```

```{sql, eval = FALSE}
-- Insert into
INSERT INTO link_copy
SELECT * FROM link
WHERE name = 'Bing';
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
INSERT INTO link_copy
SELECT * FROM link
WHERE name = 'Bing'
")
```

```{sql, eval = FALSE}
-- Select
SELECT *
FROM link_copy;

```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="learning", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT *
FROM link_copy
")
```