---
title: 'SQL: Stanford Exercises - Movie Rating Website - PART III'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Stanford's Exercises - Movie Rating - PART III

You've started a new movie-rating website, and you've been collecting data on reviewers' ratings of various movies.


### Problem 22

**Problem**: Add the reviewer Roger Ebert to your database, with an rID of 209. 

```{sql, eval = FALSE}
-- Add the reviewer Roger Ebert to your database, with an rID of 209. 
INSERT INTO Reviewer(rID, name)
VALUES (209, 'Roger Ebert')
```

```{r, echo=FALSE, message=FALSE, 1}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="StanfordSQL", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
select * from Reviewer order by rID, name
                
")
```


### Problem 23

**Problem**: Insert 5-star ratings by James Cameron for all movies in the database. Leave the review date as NULL. 

```{sql, eval = FALSE}
-- Insert 5-star ratings by James Cameron for all movies in the database. Leave the review date as NULL. 

INSERT INTO Rating  ( rID, mID, stars, ratingDate )
SELECT Reviewer.rID , Movie.mID, 5, NULL FROM Movie
LEFT OUTER JOIN Rating USING (mid)
LEFT OUTER JOIN Reviewer USING (rid)
WHERE Reviewer.name='James Cameron' 


```

```{r, echo=FALSE, message=FALSE, 2}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="StanfordSQL", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
select * from Rating where stars = 5 order by rID, mID
                
")
```


### Problem 24

**Problem**: For all movies that have an average rating of 4 stars or higher, add 25 to the release year. (Update the existing tuples; don't insert new tuples.) 

```{sql, eval = FALSE}
-- Find the titles of all movies that have no ratings. 
UPDATE movie
SET year = year + 25
WHERE mID IN (
  SELECT mID FROM (
  SELECT AVG(stars) AS astar, mID FROM Rating
  WHERE mID=rating.mID
  GROUP BY mID
  HAVING AVG(stars) >=4) d
) 
```

```{r, echo=FALSE, message=FALSE, 3}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="StanfordSQL", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
select * from Movie order by mID
                
")
```

### Problem 25

**Problem**: Remove all ratings where the movie's year is before 1970 or after 2000, and the rating is fewer than 4 stars. 

```{sql, eval = FALSE}
-- Remove all ratings where the movie's year is before 1970 or after 2000, and the rating is fewer than 4 stars. 
DELETE FROM rating
WHERE mID IN (SELECT mID FROM movie WHERE year <1970 OR year > 2000)
AND stars < 4;
```

```{r, echo=FALSE, message=FALSE}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="StanfordSQL", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
select R.rID, R.mID, R.stars, M.title, M.year from Rating R join Movie M on (R.mID = M.mID) order by R.rID, R.mID
                
")
```

