---
title: 'SQL: Stanford Exercises - Social Network - PART III'
author: "Sulman Khan"
date: "`r format(Sys.time(), ' %B %d, %Y')`"
output: 
  html_document:
    css: C:/Users/sulma/OneDrive/Desktop/Data Science/Machine Learning/gf_small_touches.css
    highlight: tango
    mathjax: default
    theme: cerulean
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
require(knitr)
options(width = 300 , scipen = 5)
options(dplyr.print_max = 200)
# options(width = 100, digits = 7)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE, 
               collapse = TRUE, tidy = FALSE,
               cache = TRUE, cache.path = '.cache/', 
               fig.align = 'left', dpi = 100, fig.path = 'figures/Introductiontotextanalytics/')
# opts_chunk$set(dev="png", 
#                dev.args=list(type="cairo"),
#                dpi=96)
```
---

## SQL: Stanford's Exercises - Social Network - PART III

You've started a new movie-rating website, and you've been collecting data on reviewers' ratings of various movies.


### Problem 15

**Problem**: Add the reviewer Roger Ebert to your database, with an rID of 209. 

```{sql, eval = FALSE}
-- It's time for the seniors to graduate. Remove all 12th graders from Highschooler.  
Delete FROM Highschooler
WHERE grade = 12;
```

```{r, echo=FALSE, message=FALSE, 1}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="StanfordSQL", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT * FROM Highschooler ORDER BY ID
                
")
```


### Problem 16

**Problem**: If two students A and B are friends, and A likes B but not vice-versa, remove the Likes tuple. 

```{sql, eval = FALSE}
-- If two students A and B are friends, and A likes B but not vice-versa, remove the Likes tuple. 

DELETE FROM Likes
WHERE ID2 IN (SELECT ID2 FROM Friend WHERE Likes.ID1 = ID1) AND
ID2 NOT IN (SELECT L.ID1 FROM Likes L WHERE Likes.ID1 = L.ID2);


```

```{r, echo=FALSE, message=FALSE, 2}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="StanfordSQL", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
SELECT H1.name, H1.grade, H2.name, H2.grade FROM Likes L, Highschooler H1, Highschooler H2 WHERE L.ID1 = H1.ID AND L.ID2 = H2.ID ORDER BY H1.name, H1.grade
                
")
```


### Problem 17

**Problem**: For all cases where A is friends with B, and B is friends with C, add a new friendship for the pair A and C. Do not add duplicate friendships, friendships that already exist, or friendships with oneself. (This one is a bit challenging; congratulations if you get it right.) 

```{sql, eval = FALSE}
-- Find the titles of all movies that have no ratings. 
INSERT INTO Friend
SELECT DISTINCT F1.ID1, F2.ID2
FROM Friend F1, Friend F2
WHERE F1.ID2 = F2.ID1 AND F1.ID1<>F2.ID2 
      AND F1.ID1 NOT IN (SELECT F3.ID1 FROM Friend F3 WHERE F3.ID2=F2.ID2);
```

```{r, echo=FALSE, message=FALSE, 3}
library(RPostgreSQL)
connPostgres = dbConnect(
  dbDriver("PostgreSQL"), dbname="StanfordSQL", 
  host="localhost", port=5432, 
  user="postgres", password="asshole"
)
DBI::dbGetQuery(connPostgres,"
select ID, name, grade, (select count(*) from Friend where id1 = H.id) from Highschooler H order by ID
                
")
```
